---
title: "RWorksheet_Palabrica#4b"
author: "Marvin Luiz Palabrica"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Using Loop Function**

**for() loop**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in

Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

**Hint** Use abs() function to get the absolute value
```{r}
matrixA <- matrix(0, nrow = 5, ncol = 5)
vectorA <- c(1, 2, 3, 4, 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixA[i, j] <- abs(i - j)
  }
}

print(matrixA)



rows <- 5

for (i in 1:rows) {
  cat(rep('"*"', i), "\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.


**Using Basic Graphics (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the data set? Show your codes and its result
```{r}
library(readr)

data <- read_csv("/cloud/project/Worksheet#4/shoe_size2.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
female_subset <- subset(data, Gender == "F")
male_subset <- subset(data, Gender == "M")

# Count observations in each subset
female_count <- nrow(female_subset)
male_count <- nrow(male_subset)

# Print counts for each gender
cat("Female:", female_count, "\n")
cat("Male:", male_count, "\n")
```
 c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
# Count the number of males and females
gender_counts <- table(data$Gender)

# Create a barplot
barplot(gender_counts, main = "Number of Males and Females",
        col = c("yellow", "green"), names.arg = c("Female", "Male"),
        ylab = "Count", xlab = "Gender", legend = TRUE)
```

5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous

a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.








