---
title: "RWorksheet_Palabrica#4a"
author: "Marvin Luiz Palabrica"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
table <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 
                10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5,
                7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 
             74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0,
             64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", 
             "M", "F", "M", "M", "M", "M", "F", "F", "M", "F",
             "F", "M", "M", "F", "M", "M", "M", "M")
)
table
```
a. Describe the data.

- The data contains measurements of 28 person which include their heights, shoe sizes, and genders. The shoe sizes range from 5.0 to 13.0, while heights vary from 59.0 to 77.0 inches.

b. Create a subset by males and females with their corresponding shoe size and height.

 What its result? Show the R scripts.
```{r}
males <- subset(table, Gender == "M", select = c(Shoe_size, Height))
females <- subset(table, Gender == "F", select = c(Shoe_size, Height))

males
females
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
MeanShoeSize <- mean(table$Shoe_size)
MeanShoeSize

meanHeight <- mean (table$Height)
meanHeight
```
d. Is there a relationship between shoe size and height? Why?
 
- Yes, there is a relationship between the shoe size and height of the respondents for the reason that the the shoesize aligns to their height.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

Consider data consisting of the names of months:
"March","April","January","November","January", "September","October","September","November","August",
"January","November","November","February","May","August, "July","December","August","August","September","November","February",
April"
```{r}
months_vector <- c("March","April","January","November","January", "September","October","September","November","August",
"January","November","November","February","May","August","July", "December","August","August","September","November","February",
"April")
 factor_months_vector <- factor (months_vector)
 factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector.
```{r}
summary(months_vector)
summary(factor_months_vector)
```
Interpret the results of both vectors. Are they both equally useful in this case?

-Yes, the are both equally useful in this case.

4. Create a vector and factor for the table below.
```{r}

factor_data <-c("East","West","North")
levels <- c(1, 4, 3)

factor_data
levels

new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)

```
5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

 View the data set. Write the R scripts and its result.
```{r}
library(readr)
import_march <- read_csv("/cloud/project/import_march.csv")
import_march
```






