---
title: "RWorksheet_Palabrica#4b"
author: "Marvin Luiz Palabrica"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**Using Loop Function**

**for() loop**

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in

Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

**Hint** Use abs() function to get the absolute value
```{r}
matrixA <- matrix(0, nrow = 5, ncol = 5)
vectorA <- c(1, 2, 3, 4, 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixA[i, j] <- abs(i - j)
  }
}

print(matrixA)



rows <- 5

for (i in 1:rows) {
  cat(rep('"*"', i), "\n")
}
```
2. Print the string “*” using for() function. The output should be the same as shown in Figure
```{r}
rows <- 5
for (i in 1:rows) {
cat(rep("''*''", i), "\n")
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input
up to 500. Use repeat and break statements. Write the R Scripts and its output.


**Using Basic Graphics (plot(),barplot(),pie(),hist())**

4. Import the dataset as shown in Figure 1 you have created previously.

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of
the data set? Show your codes and its result
```{r}
library(readr)

data <- read_csv("/cloud/project/Worksheet#4/shoe_size2.csv")
data
```
b. Create a subset for gender(female and male). How many observations are there in
Male? How about in Female? Write the R scripts and its output.
```{r}
female_subset <- subset(data, Gender == "F")
male_subset <- subset(data, Gender == "M")

# Count observations in each subset
female_count <- nrow(female_subset)
male_count <- nrow(male_subset)

# Print counts for each gender
cat("Female:", female_count, "\n")
cat("Male:", male_count, "\n")
```
 c. Create a graph for the number of males and females for Household Data. Use plot(),
chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and
its result.
```{r}
# Count the number of males and females
gender_counts <- table(data$Gender)

# Create a barplot
barplot(gender_counts, main = "Number of Males and Females",
        col = c("#FEFAE0", "#DDA15E"), names.arg = c("Female", "Male"),
        ylab = "Count", xlab = "Gender", legend = TRUE)
```

5. The monthly income of Dela Cruz family was spent on the following:
Food Electricity Savings Miscellaneous.

a. Create a piechart that will include labels in percentage.Add some colors and title of
the chart. Write the R scripts and show its output.
```{r}
categories <- c("Food", "Electricity", "Savings", "Miscellaneous")
values <- c(60, 10, 5, 25)
pie_chart <- pie(values,

labels = paste(categories, "(", values, ")", sep = ""),
col = c("#A31621", "#CED3DC", "#4E8098", "#90C2E7"),
main = "Monthly Income Distribution of Dela Cruz Family")

percentages <- round(values / sum(values) * 100, 1)
text(0, 5, paste(percentages, "%"), cex = 1.2, pos = 3)
```
6. Use the iris dataset. data(iris)
a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
data(iris)
str(iris)
```
The output of the str(iris) function shows that the iris dataset is a data frame with 150 rows and 5 columns. Four of the columns contain numeric data, which correspond to measurements of various parts of the iris flower, while the fifth column is a factor that indicates the flower's species.

b. Create an R object that will contain the mean of the sepal.length, sepal.width,petal.length,and
petal.width. What is the R script and its result?
```{r}
meanValues <- colMeans(iris[, 1:4])
meanValues
```
c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and
its result.
```{r}
data(iris)
speciesCounts <- table(iris$Species)
pie(speciesCounts,
main = "Species Distribution in Iris Dataset",
col = c("#F1E4F3", "#F4BBD3", "#F686BD"), labels = paste(names(speciesCounts), "\n", speciesCounts))
legend("topright", legend = names(speciesCounts), fill = c("#F1E4F3", "#F4BBD3", "#F686BD"))
```
d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r}
setosa_data <- iris[iris$Species == "setosa", ]
versicolor_data <- iris[iris$Species == "versicolor", ]
virginica_data <- iris[iris$Species == "virginica", ]
last_six_setosa_data <- tail(setosa_data, 6)
last_six_versicolor_data <- tail(versicolor_data, 6)
last_six_virginica_data <- tail(virginica_data, 6)
last_six_setosa_data
last_six_versicolor_data
last_six_virginica_data
```
e. Create a scatterplot of the sepal.length and sepal.width using the different species(setosa,versicolor,virginica).

Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width,
main = "Iris Dataset",
sub = "Sepal Width and Length",
xlab = "Sepal Length",
ylab = "Sepal Width",
pch = 19,
col = iris$Species)
legend("topright", legend = levels(iris$Species),
col = 1:3, pch = 19)
```
f.Interpret the result.

The pie chart displays the distribution of iris species in the dataset, with each slice representing the proportion of each species. This helps identify the most common species. The subsets show the last six entries for each species, providing detailed measurements of sepal and petal dimensions, which helps compare how these measurements vary across species. The scatterplot reveals the relationship between sepal length and width. If setosa forms a cluster at lower values, it indicates that this species generally has smaller flowers. On the other hand, if versicolor and virginica overlap, it suggests that their flower sizes are quite similar, making it difficult to distinguish between them based on these measurements.

**BASIC CLEANING AND TRANSFORMATION OF OBJECTS**

7. Import the alexa-file.xlsx. Check on the variations. Notice that there are ex- tra whitespaces among
black variants (Black Dot, Black Plus, Black Show, Black Spot). Also on the white variants (White Dot, White Plus, White Show, White Spot).

a. Rename the white and black variants by using gsub() function.
```{r}
library(readxl)
alexa_data <- read_excel("/cloud/project/Worksheet#4/alexa_file.xlsx")
unique(alexa_data$variation)
```

```{r}
alexa_data$variation <- gsub("Black Dot", "BlackDot", alexa_data$variation)
alexa_data$variation <- gsub("Black Plus", "BlackPlus", alexa_data$variation)
alexa_data$variation <- gsub("Black Show", "BlackShow", alexa_data$variation)
alexa_data$variation <- gsub("Black Spot", "BlackSpot", alexa_data$variation)
alexa_data$variation <- gsub("White Dot", "WhiteDot", alexa_data$variation)
alexa_data$variation <- gsub("White Plus", "WhitePlus", alexa_data$variation)
alexa_data$variation <- gsub("White Show", "WhiteShow", alexa_data$variation)
alexa_data$variation <- gsub("White Spot", "WhiteSpot", alexa_data$variation)
unique(alexa_data$variation)
```
b. Get the total number of each variations and save it into another object. Save the object as variations. RData. Write the R scripts. What is its result? Hint: Use the dplyr package. Make sure to install it before loading the package.
```{r}
library(dplyr)
variations <- alexa_data %>%
count(variation)
print(variations)
```
**SAMPLE OUTPUT**
c. From the variations.RData, create a barplot(). Complete the details of the chart which include the
title, color, labels of each bar.
```{r}
library(dplyr)
variations$variation <- gsub(" +", " ", variations$variation)
variations$variation <- trimws(variations$variation)
bardata <- variations$n
barnames <- variations$variation
barplot(
bardata,
main = "Counts of Variations",
col = rainbow(length(bardata)),
names.arg = barnames,
xlab = "Variations",
ylab = "Counts",
las = 2,
border = "#53131e"
)
text(
x = seq_along(bardata),
y = bardata + max(bardata) * 0.05,
labels = bardata,
pos = 3,
cex = 1,
col = "#53131e"
)
```
d. Create a barplot() for the black and white variations. Plot it in 1 frame, side by side. Complete the
details of the chart.
```{r}
library(ggplot2)
library(dplyr)

variations$variation <- gsub(" +", " ", variations$variation)
variations$variation <- trimws(variations$variation)
bwvariations <- variations %>%
filter(grepl("Black|White", variation))
bardata <- as.matrix(bwvariations$n)
barnames <- bwvariations$variation
barplot(
bardata,
beside = TRUE,
main = "Counts of Black and White Variations",
col = c("#FA7921", "#FE9920", "#B9A44C", "#566E3D"),
names.arg = barnames,
xlab = "Variations",
ylab = "Counts",
las = 2,
border = "black"
)
text(x = seq_along(bardata), y = bardata, labels = bardata, pos = 3, cex = 0.8, col = "black")
```















