---
title: "RWorksheet_Palabrica#4a"
author: "Marvin Luiz Palabrica"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.
```{r}
table <- data.frame(
  Shoe_size = c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 
                10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5,
                7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0),
  
  Height = c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 
             74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0,
             64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0),
  
  Gender = c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", 
             "M", "F", "M", "M", "M", "M", "F", "F", "M", "F",
             "F", "M", "M", "F", "M", "M", "M", "M")
)
table
```
a. Describe the data.

- The data contains measurements of 28 person which include their heights, shoe sizes, and genders. The shoe sizes range from 5.0 to 13.0, while heights vary from 59.0 to 77.0 inches.

b. Create a subset by males and females with their corresponding shoe size and height.

 What its result? Show the R scripts.
```{r}
males <- subset(table, Gender == "M", select = c(Shoe_size, Height))
females <- subset(table, Gender == "F", select = c(Shoe_size, Height))

males
females
```
c. Find the mean of shoe size and height of the respondents. Write the R scripts and its result.
```{r}
MeanShoeSize <- mean(table$Shoe_size)
MeanShoeSize

meanHeight <- mean (table$Height)
meanHeight
```
d. Is there a relationship between shoe size and height? Why?
 
- Yes, there is a relationship between the shoe size and height of the respondents for the reason that the the shoesize aligns to their height.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

Consider data consisting of the names of months:
"March","April","January","November","January", "September","October","September","November","August",
"January","November","November","February","May","August, "July","December","August","August","September","November","February",
April"
```{r}
months_vector <- c("March","April","January","November","January", "September","October","September","November","August",
"January","November","November","February","May","August","July", "December","August","August","September","November","February",
"April")
 factor_months_vector <- factor (months_vector)
 factor_months_vector
```
3. Then check the summary() of the months_vector and factor_months_vector.
```{r}
summary(months_vector)
summary(factor_months_vector)
```
Interpret the results of both vectors. Are they both equally useful in this case?

-Yes, the are both equally useful in this case.

4. Create a vector and factor for the table below.
```{r}

factor_data <-c("East","West","North")
levels <- c(1, 4, 3)

factor_data
levels

new_order_data <- factor(factor_data,levels = c("East","West","North"))
print(new_order_data)

```
5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

 View the data set. Write the R scripts and its result.
```{r}
library(readr)
import_march <- read_csv("/cloud/project/import_march.csv")
import_march
```
6. Full Search

Exhaustive search is a methodology for finding an answer by exploring all possible cases.
When trying to find a desired number in a set of given numbers, the method of finding the corresponding number by checking all elements in the set one by one can be called an exhaustive search. Implement an exhaustive search function that meets the input/output conditions below.

a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string "The number selected is beyond the range of 1 to 50". If number 20 is inputted by the User, it will have to display "TRUE", otherwise display the input number.
```{r}
ExhaustiveSearch <- function(selectedNumber) {
  if (selectedNumber < 1 || selectedNumber > 50) {
    return("The number selected is beyond the range of 1 to 50")
  } else if (selectedNumber == 20) {
    return("TRUE")
  } else {
    return(as.character(selectedNumber))
  }
}

cat("Please select a number between 1 and 50:\n")
userInput <- as.numeric(readLines(con = "stdin", n = 1))

result <- ExhaustiveSearch(userInput)
cat(result, "\n")

```
7. Change

At ISATU University's traditional cafeteria, snacks can only be purchased with bills. A long-standing rule at the concession stand is that snacks must be purchased with as few coins as possible. There are three types of bills: 50 pesos, 100 pesos, 200 pesos, 500 pesos, 1000 pesos.

a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack.

Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.
7. Change

At ISATU University's traditional cafeteria, snacks can only be purchased with bills. A long-standing rule at the concession stand is that snacks must be purchased with as few coins as possible. There are three types of bills: 50 pesos, 100 pesos, 200 pesos, 500 pesos, 1000 pesos.

a. Write a function that prints the minimum number of bills that must be paid, given the price of the snack.

Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.
```{r}
minimum_bills <- function(price) {
  type_bills <- c(1000, 500, 200, 100, 50)
  count <- 0
  
  for (bill in type_bills) {
    if (price >= bill) {
      count <- count + (price %/% bill)
      price <- price %% bill
    }
  }
  
  print(paste("Minimum number of bills needed:", count))
}

minimum_bills(200)



```

