---
title: "RWorksheet_Palabrica#3b"
author: "Marvin Luiz Palabrica"
date: "2024-09-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1. Create a data frame using the table below.

a. Write the codes.
```{r}
data <- data.frame(
Respondents = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20),
Sex = c(2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 2, 2, 1, 2),
FathersOccupation = c(1, 3, 3, 3, 1, 2, 3, 1, 1, 1, 3, 2, 1, 3, 3, 1, 3, 1, 2, 1),
PersonsAtHome = c(5, 7, 3, 8, 5, 9, 6, 7, 8, 4, 7, 5, 4, 7, 8, 8, 3, 11, 7, 6),
SiblingsAtSchool = c(6, 4, 4, 1, 2, 1, 5, 3, 1, 2, 3, 2, 5, 5, 2, 1, 2, 5, 3, 2),
TypeofHouses = c(1, 2, 3, 1, 1, 3, 3, 1, 2, 3, 2, 3, 2, 2, 3, 3, 3, 3, 3, 2)
)

colnames(data) <- c("Respondents", "Sex", "Fathers Occupation", "Persons At Home", "Siblings At School", "TypeofHouses")
data

legend <- list(
Sex = c("Male" = 1, "Female" = 2),
FathersOccupation = c("Farmer" = 1, "Driver" = 2, "Others" = 3),
TypeofHouses = c("Wood" = 1, "Semi-concrete" = 2, "Concrete" = 3)
)
legend
```
b. Describe the data. Get the structure or the summary of the data    

The dataset includes responses from 20 participants, with variables like sex (coded as 1 and 2), fathers' occupations (values from 1 to 3), the number of individuals living at home (ranging from 3 to 11), siblings attending school (from 1 to 6), and types of housing (values between 1 and 3).

```{r}
str(data)
summary(data)
```

    
c. Is the mean number of siblings attending is 5?

NO
```{r}
meansiblings <- mean(data$SiblingsAtSchool)

```
d. Extract the 1st two rows and then all the columns using the subsetting functions.
Write the codes and its output.    
```{r}
subset_data <- data[1:2, ]
subset_data
```
e. Extract 3rd and 5th row with 2nd and 4th column. Write the codes and its result.
```{r}
subset_data <- data[c(3, 5), c(2, 4)]
subset_data
```
f. Select the variable types of houses then store the vector that results as types_houses. Write the codes.
```{r}
types_houses <- data$TypeofHouses
types_houses
```
g. Select only all Males respondent that their father occupation was farmer. Write the codes and its output.   
```{r}
malesfarmers <- data[data$Sex == 1 & data$FathersOccupation == 1, ]
malesfarmers

```
h. Select only all females respondent that have greater than or equal to 5 number of siblings attending school. Write the codes and its outputs.
```{r}
femalessiblings <- data[data$Sex == 2 & data$SiblingsAtSchool >= 5, ]
femalessiblings
```
2. Write a R program to create an empty data frame. Using the following codes:
```{r}
df = data.frame(Ints=integer(),

Doubles=double(), Characters=character(),
Logicals=logical(),
Factors=factor(),
stringsAsFactors=FALSE)

print("Structure of the empty dataframe:")
print(str(df))
```
a. Describe the results.


3. Create a .csv file of this. Save it as HouseholdData.csv

a. Import the csv file into the R environment. Write the codes.
```{r}
data1 = read.csv("/cloud/project/HouseholdData.csv")
data1
```
b. Convert the Sex into factor using factor() function and change it into integer.[Legend: Male = 1 and Female = 2]. Write the R codes and its output.
```{r}
data1$Sex <- factor(data1$Sex, levels = c("Male", "Female"), labels = c(1, 2))
data1$Sex
```
c. Convert the Type of Houses into factor and change it into integer. [Legend: Wood = 1; Congrete = 2; Semi-Congrete = 3]. Write the R codes and its output.
```{r}
data1$TypeofHouse <- factor(data1$TypeofHouse,
                            
levels = c("Wood", "Congrete", "Semi-congrete"),
labels = c(1, 2, 3))

data1$TypeofHouse
```
d. On father’s occupation, factor it as Farmer = 1; Driver = 2; and Others = 3. What is the R code and its output?
```{r}
data1$FathersOccupation <- factor(data1$FathersOccupation,
levels = c(1, 2, 3),
labels = c("Farmer", "Driver", "Others"))
data1$FathersOccupation
```
e. Select only all females respondent that has a father whose occupation is driver. Write the codes and its output.
```{r}
femaledrivers <- data1[data1$Sex == 2 & data1$FathersOccupation == 2, ]
femaledrivers
```
f. Select the respondents that have greater than or equal to 5 number of siblings attending school. Write the codes and its output.
```{r}
siblings_filter <- data1[data1$SiblingsAtSchool >= 5, ]
siblings_filter

```
4. Interpret the graph.

The graph shows a sentiment analysis of tweets from July 14 to July 21, 2020, grouping them as negative, neutral, or positive. Negative tweets were the most common on most days, especially on July 15 and July 21, with July 15 reaching about 4000 negative tweets. In contrast, July 18 and July 20 had fewer tweets overall, particularly in the negative category. Positive tweets stayed fairly steady throughout the week but were always fewer than negative ones, showing that negative sentiment was the strongest during this period.





